---
title: "Predicting ecosystem responses by data-driven reciprocal modelling"
subtitle: "Supplementary Material"
author: "Florian Schneider, Christopher Poeplau and Axel Don"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document 
papersize: a4
bibliography: literature.bib
csl: global-change-biology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F)
knitr::opts_chunk$set(dev = 'pdf')
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(here)
library(kableExtra)
library(sf)
library(latex2exp)
library(patchwork)

goodData=readRDS(here("data","derived",
                      "goodData.rds")) %>% 
  mutate(obs=OCstock)

europe=st_as_sf(rnaturalearth::countries110) %>%
  filter(region_un=="Europe"&!(name%in%c('Russia',"Ukraine","Belarus","Moldova"))) %>%
  st_crop(xmin=-10,
          xmax=45,
          ymin=30,
          ymax=73)
```

## Overview of explanatory variables
Table 1. Variables used in the LUCAS case study. Sources: lucas [@Jones2020; @Orgiazzi2018], worldclim [@Fick2017; @Harris2014], ef_lus_allcrops [@ef_lus_allcrops], ef_lus_main [@ef_lus_main], ef_m_farmleg [@ef_m_farmleg], aei_fm_ms [@aei_fm_ms], ef_lsk_main [@ef_lsk_main], aei_ps_inp [@aei_ps_inp], aei_fm_usefert [@aei_fm_usefert], ef_mp_irri [@ef_mp_irri], opendatascience [@opendatascience].
```{r table}
options(knitr.kable.NA = '')
vars=readxl::read_excel(here("vars.xlsx")) %>%
  filter(class0!="0info",
         modelEffect==1) %>% 
  ungroup() %>% 
  arrange(desc(class0)) %>% 
  mutate(modelSOC=ifelse(modelSOC==1,"x",""),
         modelEffect=ifelse(modelEffect==1,"x",""),
         nameFull=str_trunc(nameFull,50,"right"),
         type=case_when(name=="Soil_Stones"~"F(4)",
                        type=="numeric"~"C",
                        T~paste0("F(",nlevs,")")))

temp=vars %>% 
  mutate(row=row_number(),
         class0=ifelse(class0=="target","soil",class0)) %>% 
  group_by(class0) %>% 
  summarise(min=min(row),
            max=max(row))

vars %>% 
  select(nameFull,type,unit,modelSOC,modelEffect,source) %>% 
  kbl(col.names = c("Name",
                    "Type*",
                    "Unit$^{\\dagger}$",
                    "SOC stock",
                    "effect size",
                    "Source"),
      longtable=T,
      booktabs=T,
      escape=F) %>% 
  kable_paper(font_size=6) %>% 
  add_header_above(c(" "=1," "=1," "=1,"Explanatory variable for"=2," "=1)) %>% 
  footnote(symbol =c("C = Continuous variable; F(L) = Categorical variable with L levels",
                     "UAA = Utilizable (total) agricultural area")) %>% 
  pack_rows("Pedosphere",temp$min[temp$class0=="soil"],temp$max[temp$class0=="soil"]) %>%
  pack_rows("Geosphere",temp$min[temp$class0=="geo"],temp$max[temp$class0=="geo"]) %>%
  pack_rows("Atmosphere",temp$min[temp$class0=="clim"],temp$max[temp$class0=="clim"]) %>%
  pack_rows("Anthroposphere",temp$min[temp$class0=="anthro"],temp$max[temp$class0=="anthro"])

```

## Maps
```{r mapSoil, message=F, warning=F, fig.width=7, fig.height = 11, out.width = "\\textwidth", results='asis'}
# Soil variables
dat=vars %>%
  filter(type=="C") %>%
  select(name,nameFull,unit)

# Create 4 row x 5 column grid
grid=expand.grid(row=LETTERS[seq(from=1,to=4)],
                 col=1:4) %>%
  arrange(row)

dat=dat %>%
  mutate(tot=row_number(),
         page=ceiling(row_number()/nrow(grid))) %>%
  group_by(page) %>%
  mutate(temp=row_number(),
         pos=paste0(grid$row,grid$col)[temp]) %>%
  ungroup() %>%
  select(-temp)

for(n in 1:max(dat$page)){
  for(m in 1:nrow(grid)){

    item=paste0(grid$row,grid$col)[m]
    case=dat %>%
       filter(page==n,
              pos==item)

    if(nrow(case)==0) {

      assign(item,
             plot_spacer())

    } else {

  sub=goodData %>%
    select(Point_ID,lon,lat,one_of(case %>% pull(name))) %>%
    drop_na() %>%
    st_as_sf(coords = c("lon", "lat"),
             crs = 4326)
  names(sub)[2]="target"

    bks2 = cartography::getBreaks(v = sub$target, nclass = 5,  method = "fisher-jenks")
    bks2.round=bks2
    if(median(bks2)>.001){
      bks2.round=round(bks2,4)
    }
    if(median(bks2)>.01){
      bks2.round=round(bks2,3)
    }
    if(median(bks2)>.1){
      bks2.round=round(bks2,2)
    }
    if(median(bks2)>1){
      bks2.round=round(bks2,1)
    }
    if(median(bks2)>10){
      bks2.round=round(bks2,0)
    }
    if(median(bks2)>100){
      bks2.round=round(bks2,0)
    }
    pal = colorspace::sequential_hcl(5,"YlGnBu",rev=T)

hist=ggplot(sub,
            aes(target)) +
  geom_histogram(breaks=bks2,
                 fill=pal) +
  scale_x_continuous(breaks=bks2,
                     labels=bks2.round,
                     guide=guide_axis(n.dodge=2))+
  scale_y_continuous(expand = c(0,0))+
  theme_void()+
  theme(axis.ticks.x=element_line(colour="black",size=.1),
        axis.ticks.length=unit(2, "pt"),
        axis.text.x=element_text(size=6),
        plot.caption=element_text(size=6),
        panel.grid = element_blank(),
        panel.border = element_blank())
if(!is.na(case$unit)){
  hist=hist+
    labs(caption=TeX(case$unit))
  }

plot=ggplot() +
  labs(title=paste(strwrap(case$nameFull,
                           width = 30),
                   collapse = "\n"))+
  geom_sf(data = europe,
          fill = "white",
          size = .2) +
  geom_sf(data = sub %>%
            mutate(target=Hmisc::cut2(target,cuts=bks2)),
          aes(color = target),
          size = .001)+
  scale_color_manual(values=pal) +
  theme_void(base_size =6)+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position="none")

plot=plot/(plot_spacer()+hist+plot_spacer()+plot_layout(widths=c(.1,1,.1)))+plot_layout(heights=c(1.5,.1))
assign(item,plot)
       }
  }
 page=(A1|A2|A3|A4)/
   (B1|B2|B3|B4)/
   (C1|C2|C3|C4)/
   (D1|D2|D3|D4)
 # assign(paste0("page",n),page)
 print(page)
 cat("\n\n\\pagebreak\n")
 }
# page

```

\newpage
## SOC data of alkaline soils
In the present study, sites with CaCa$_3$ contents above 5\% were omitted. The reason for this was high scatter in the C:N ratio of alkaline soils, indicating substantial analytic uncertainties in SOC measurements of soil with inorganic carbon (see Figure below).  

```{r caco3, echo=FALSE, message=F, warning=F,fig.height = 3, fig.width = 4.5, unit = "cm", fig.align = "center"}
readRDS(here("data","derived",
                      "rawData.rds")) %>% 
    ggplot(aes(pH_H2O,OC/N,
             col=CaCO3<5))+
  geom_point(shape=1,alpha=.2,size=.3)+
  colorspace::scale_color_discrete_qualitative(palette = "Dark 3")+
  scale_y_log10()+
  labs(x=expression('pH'["H2O"]),
       y="C:N ratio",
       col=expression('CaCO'[3]~'< 5 %'))+
  guides(colour = guide_legend(override.aes = list(size=3,alpha=1)))+
  theme_bw()+
  theme(panel.grid = element_blank())


```



\newpage

## References
\vspace{-.55cm}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}

\noindent

<div id="refs"></div>

```{r refmgr references, results="asis", echo=FALSE}
# PrintBibliography(bib) 
```